Distance Sampling in \proglang{R}

### 2016-01-18 paper.Rmd
Updated abstract

---
author:
  - name: David L Miller
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: dave@ninepointeightone.net
    url: http://converged.yt
  - name: Eric Rexstad
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: Eric.Rexstad@st-andrews.ac.uk
  - name: Len Thomas
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: len.thomas@st-andrews.ac.uk
  - name: Laura Marshall
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: lhm@st-andrews.ac.uk
  - name: Jeffrey L Laake
    affiliation: NOAA
    address: >
      National Marine Mammal Laboratory
      Alaska Fisheries Science Center
      7600 Sand Point Way N.E., Seattle, WA 98115, USA
    email: Jeff.Laake@noaa.gov
title:
  formatted: "Distance Sampling in \\proglang{R}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Distance Sampling in R"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  (LT: I have gone through putting comments to myself about things I need to come back and edit.  In the end, I have written them within brackets in the text because I couldn't get the todo package to work, because square brackets are interpreted as markdown.  In general, we might have been better to have used .Rnw but och well.  Anyway DLM or others, feel free to comment on my comments!  (Sorry if comments seem a bit negative -- I'd rather have had time to do the editing on this round so you can see my suggested text -- but it always takes me ages to do this bit and I had limited time right now -- I'll come back to it!...))
  (LT: Need to define abundance; need to introduce concept of density also.) The estimation of the abundance of a biological population is an essential part of applied ecology and wildlife management. (LT: This sentence not strictly accurate, probably needs some editing.) To ensure that abundance estimates are accurate, we must account for uncertain detectability of plants and animals; distance sampling is a popular method for doing this. Here we show how models for detectabilty can be fitted using the \proglang{R} package \pkg{Distance}. (LT: Is it actually helpful to use the jargon like Horvitz-Thompson-like in the abstract?  Almost no-one will know what we mean.) We also show how \pkg{Distance} can be used to calculate Horvitz-Thompson-like estimates of abundance, its interface to other packages to provide sophisticated analyses and a migration pathway from previous software.
keywords:
  # at least one keyword must be supplied
  formatted: [distance sampling, abundance estimation, line transects, point transects, detection functions, Horvitz-Thompson, "\\proglang{R}", Distance]
  plain:     [distance sampling, abundance estimation, line transects, point transects, detection functions, Horvitz-Thompson, R, Distance]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage[colorinlistoftodos, color=green!40, textwidth=60, textsize=small]{todonotes}

>>>>

---
author:
  - name: David L Miller
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: dave@ninepointeightone.net
    url: http://converged.yt
  - name: Eric Rexstad
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: Eric.Rexstad@st-andrews.ac.uk
  - name: Len Thomas
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: len.thomas@st-andrews.ac.uk
  - name: Laura Marshall
    affiliation: University of St Andrews
    address: >
      Centre for Research into Ecological and Environmental Modelling,
      The Observatory,
      St Andrews, Fife KY16 9LZ, Scotland
    email: lhm@st-andrews.ac.uk
  - name: Jeffrey L Laake
    affiliation: NOAA
    address: >
      National Marine Mammal Laboratory
      Alaska Fisheries Science Center
      7600 Sand Point Way N.E., Seattle, WA 98115, USA
    email: Jeff.Laake@noaa.gov
title:
  formatted: "Distance Sampling in \\proglang{R}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Distance Sampling in R"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >

  The study of plants or animals often involves the estimation of population size and/or their spatial distribution. These characteristics are important considerations in wildlife management. We introduce the \proglang{R} package \pkg{Distance} that uses distance sampling data to estimate populations size (abundance) and can be used in conjunction with another \proglang{R} package \pkg{dsm} to estimate spatial distribution. We describe example analyses using this package as well demonstrating the links with other \proglang{R} packages offering specialist solutions to specific distance sampling challenges. We demonstrate how this package provides a migration pathway from previous software, thereby allowing us to deliver cutting edge methods to the users more quickly.

keywords:
  # at least one keyword must be supplied
  formatted: [distance sampling, abundance estimation, line transects, point transects, detection functions, Horvitz-Thompson, "\\proglang{R}", Distance]
  plain:     [distance sampling, abundance estimation, line transects, point transects, detection functions, Horvitz-Thompson, R, Distance]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{longtable}
  \usepackage{booktabs}
  \usepackage[colorinlistoftodos, color=green!40, textwidth=60, textsize=small]{todonotes}
